\documentclass[12pt, a4paper]{report}
\usepackage{epsfig}
\usepackage{subfigure}
%\usepackage{amscd}
\usepackage{amssymb}
\usepackage{graphicx}
%\usepackage{amscd}
\usepackage{amssymb}
\usepackage{amsthm, amsmath}
\usepackage{amsbsy}
\usepackage[usenames]{color}
%\usepackage{listings}

\title{Descriptive Statistics}


\begin{document}


\section*{Skewness and Kurtosis}

A fundamental task in many statistical analyses is to characterize the location and variability of a data set. A further characterization of the data includes skewness and kurtosis. 

\begin{itemize}
\item \textbf{Skewness} is a measure of symmetry, or more precisely, the lack of symmetry. A distribution, or data set, is symmetric if it looks the same to the left and right of the center point.

\item \textbf{Kurtosis} is a measure of whether the data are peaked or flat relative to a normal distribution. That is, data sets with high kurtosis tend to have a distinct peak near the mean, decline rather rapidly, and have heavy tails. Data sets with low kurtosis tend to have a flat top near the mean rather than a sharp peak. A uniform distribution would be the extreme case.

\item The histogram is a reasonably effective graphical technique for showing both the skewness and kurtosis of data set.

\end{itemize}

%-------------------------------------------------------------------------%
\frame{
\frametitle{Symmetric and Skewed Distributions }
\begin{itemize}
\item  A data set is said to be \textbf{\emph{symmetric}} when data values are distributed in the same way above and below the middle of the sample.
\item Typically , but not necessarily, distributions are considered to symmetric because the data set does not contain any outliers. (Potentially a symmetric data set may contain outliers on either side of the mean. )
\item When data sets are symmetrically distributed, the sample mean and median have values close to each other.
\item Distributions are considered to \textbf{\emph{skewed}} when the data set contains an outlier, or cluster of outliers, distributed away from the main cluster of items, such as in our example.
\item When data sets have skewed distribution, the sample mean and median have values different to each other.
\end{itemize}
}
%-------------------------------------------------------------------------%
\frame{
\frametitle{Symmetric and Skewed Distributions }
\begin{itemize}
\item When a data set is \textbf{\emph{symmetric}}, the best measure of centrality is the sample mean.
\item As the variance is computed using the sample mean, it is considered the best measure of dispersion for symmetrically distributed data. \bigskip
\item When a data set is \textbf{\emph{skewed}}, the best measure of centrality is the median.
\item The best measure of dispersion for data with skewed distribution is the IQR.
\end{itemize}
}

\end{document}

