Question 1 

Combinations and Permutations: From first principles, compute the following:
1.	5C2
2.	4C0
3.	4C4
4.	10C1
5.	10C0

Question 2

Six horses are entered in a race. Larry picks two of the horses at random, and bets on them. 
How many ways are there of picking two horses at random from six?

Question 3

Suppose a fair coin is tossed six times. The number of heads which can occur with their respective
probabilities are as follows:

xi	0	1	2	3	4	5	6
p(xi)	1/64	6/64	15/64	20/64	15/64	6/64	1/64

a)	Compute the expected value (i.e. expected number of heads).
b)	Compute the variance of the number of heads.
Question 4
A player tosses two fair coins. He wins $2 if two heads occur, and $1 if one head occurs. On the other hand, he loses $3 if no heads occur. 

Find the expected value E(X) of the game. Is the game fair? 

(The game is fair, favourable, or unfavourable to the player if E(X) = 0,E(X) > 0 or E(X) < 0 respectively)

%---------------------------------------------------------------------------------%
Question 5
Consider a binomial experiment where the number of independent trials is 10, and the probability of success is 0.25. Let X denote the number of successes in the 10 trials.
A)	Write out the sample space of X.
B)	For the each of the following events, write out the sample points:
i)	the number of successes is less than or equal to 3,
ii)	the number of successes is greater than 4, 
iii)	the number of successes is greater than 8, 
iv)	the number of successes is between 4 and 7 inclusive.
%---------------------------------------------------------------------------------%Question 6
By considering the relevant sample points from the sample space in the previous question, show that 
1)	P(4 ≤X ≤8) = P(X ≤8) – P(X ≤3 ) = P(X  4) – P(X  9)
2)	P(X=4) = P(X  4) – P(X  5 )
3)	P(X≥7) = 1- P(X ≤6)
(i.e. show how the sample space can be partitioned accordingly)
Question 7: Binomial Distribution: 
Suppose that a student is taking a multiple-choice exam in which each question has four choices.
Suppose that she has no knowledge of the correct answer to any of the questions. Furthermore suppose that she selects one of the possible choices at random as her answer.
A.	lf there are five muItiple—choice questions on the exam, what is the probability that she will answer four questions correctly.
B.	What is the probability that she will answer none of the questions correctly?
C.	What is the probability that she will answer at least two questions correctly?
Question 8
A company owns 400 laptops. Each laptop has a 2% probability of not working. You randomly select 20 laptops for your salespeople.
A. What distribution do you feel is appropriate here. Justify your answer. 
B. What is the probability that they will all work?
C. What is the probability that only one will be broken?
D. What is the probability that at least two will be broken? 
Question 9:  Poisson Distribution
Cars enter a car wash at an average rate of 4 per half hour. 
1)	Compute the probability that the 3 cars arrive in a half hour period.
2)	Compute the probability that the 3 cars arrive in a 15 minute period.

