\begin{document}
%--------------------------%
\begin{frame}
\frametitle{The Weibull Distribution}

\Large



\end{frame}
%--------------------------%

%--------------------------%
\begin{frame}
\frametitle{The Weibull Distribution}

\Large

\begin{itemize}
\item The two-parameter Weibull distribution is the most widely used distribution for life data analysis. Apart from the 2-parameter Weibull distribution, the 3-parameter and the 1-parameter Weibull distribution are often used for detailed analysis.
\item 
The 2-parameter Weibull cumulative distribution function (CDF), has the explicit equation:
\end{itemize

\end{frame}
%--------------------------%
%--------------------------%
\begin{frame}
\frametitle{The Weibull Distribution}

\Large

\begin{itemize}
\item F(t) = Probability of failure at time t;
\item t = time, cycles, miles, or any appropriate parameter;
\item = characteristic life or scale parameter;
\item = slope or shape parameter.
\end{itemize}


Note that for no other, apart from the Exponential , such an explicit equation is available. Below the F(t) equations for the


 
\end{frame}
%--------------------------%
%--------------------------%
\begin{frame}
\frametitle{Weibull Distribution}

Normal and the Log-Normal distributions.

The Weibull Probability Density Function (PDF) is:


\end{frame}
%--------------------------%
%--------------------------%
\begin{frame}
\frametitle{Weibull Distribution}

Weibull Parameters


The two parameters of the 2-Parameter Weibull distribution are:
 , the slope or shape parameter, and
 , the characteristic life or scale parameter.
\end{frame}
%--------------------------%
%--------------------------%
\begin{frame}
Shape parameter
BETA () shows how the failure rate develops in time. Failure-modes related with Infant Mortality, Random, or Wear-Out have significant different Beta values. is named the shape parameter as it determines which member of the Weibull family of distributions 
is most appropriate. 
Different members have 
different shaped probability 
density functions (PDF
\end{frame}
%--------------------------%
\end{document}

