\documentclass{beamer}

\usepackage{default}

\begin{document}
	
	
	Exam : Wednesday March 16th at 9pm
	
	45 minutes long
	Worth 15% of the overall module grade
	No calculators allowed.
	15 multiple choice questions
	
	
	Qualitative Data
	Quantitative Data
	
	Discrete v Continuous
	
\end{frame}
%============================================================ %
\begin{frame}
	An experiment is
	
	An observational study is
	
	This is an experiment because the outcomes are a response to conditions set by the study.
	
\end{frame}
%============================================================ %
\begin{frame} 
	
	Ratio v Interval
	
	The difference between ratio and interval is most evident in how a value of zero is treated.
	
	For Ratio values – a value of zero means the complete absence.
	For interval values
	
	Consider that 0 degrees centigrade is 32 degrees Fahrenheit.
	
	ordinal
	nominal (Categorical)
	
	For example consider a variable”origin” where the outcomes are Limerick /Rest of Munster/Rest of Ireland/Rest of World.
	
	ordinal – contains a hierarchy
	
	Discrete:
	
	Number of Children in a class
	Number of faulty components per batch
	
	Continuous
	
	Length of time between cars arriving at a petrol station pump.
	
	
	It  is not possible have 24.5 children in any class, but it is possible to have an average of 24.5 per class.
\end{frame}
%============================================================ %
\begin{frame}
	
	Definitions
	Population
	Sample
	Statistic: value based on a sample. Commonly used as an estimate for the parameter value.
	
	Parameter value
	Sampling Frame:
	Databases
	
\end{frame}
%============================================================ %
\begin{frame}
	
	Section 2 Graphical methods
	
	Pie Charts – hard to make out what is going on.
	Some aesthetic use i.e. breaking up a long page of text.

\end{frame}
%============================================================ %
\section{Probability}
\begin{frame}
	
\begin{itemize}
\item Law of total probability
	
\item Independent events
	
\item Conditional probability
\end{itemize}	

\end{frame}
%============================================================ %
\begin{frame}
	
	Given that event B has already occurred ,   what is the probability of event A
	
	P( A | B)  “probability of A given B”
	
	P( A and B) “ probability of A and B”
	
	\textbf{Complement event}\\
	
	What is the probability of A not happening
	What is the probability of outcomes not included in A.
	
\end{frame}
%============================================================ %
\begin{frame}
	
	Section 4: Probability Distributions
	
	Expected value of the outcome
	
	E[X]
	
	By definition the expected value is the value that 50% of the the outcomes are greater than.By extension 50% of values are less than the Expected value.
	\[
	P( X \geq E[X])  = 0.50 \]
	\[ P(X \leq E[X]) = 0.50\]
\end{frame}
%============================================================ %
\section{Binomial Distribution}
\begin{frame}	

	Number of independent trials
	
	A coin is tossed eight times. – the number of trials is therefore 8.
	
	A group of people or a batch of items can also be considered as a series of independent trials.
\end{frame}
%============================================================ %
\begin{frame} 
	Probability of a success
	
	A “success” is dependent on how the question is framed, or what is being estimated.

\end{frame}
%============================================================ %
\begin{frame}	
	Continuous Uniform distribution
	
\begin{description}
\item[L] lower bound of an interval
\item[U] upper bound of an interval
\end{description}
	Probability of an outcome being between lower bound L and upper bound U
	\[P( L \leq X \leq U)  =  { U - L \over  b – a }\]
\end{frame}
%============================================================ %
\begin{frame} 
	
	Reminder
	
	" $\leq$" is less than or equal to
	
	" $\geq$" is greater than or equal to
	
	
	\[L \leq X \leq U\]
	
	simply states that X is between L and U inclusively.
	
	("inclusively" mean that X could be exactly L or U also, although the probability of this is extremely low)
	
\end{frame}
%============================================================ %
\begin{frame} 
	The probability density function is given as
	
	\[f(x) = {1 \over b-a} for a \leq x \leq b\]
	
	For any value "c" between the minimum value a and the maximum value b
	
	\[P(X \ geq c) = {b-c \over b-a}\]
	
	here b is the upper bound while c is the lower bound
	
	
	\[P(X \ leq c) = {c-a \over b-a}\]
	
	here c is the upper bound while a is the lower bound
	
\end{frame}
%============================================================ %
\begin{frame}
	
	Exercises on Murdoch Barnes table 6
	
	Find the following
	
	$e^{-0.6}$
	
	$e^{-1.4}$
	
	$e^{-3.2}$
	
\end{frame}
\end{document}
